---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fars.functions

<!-- badges: start -->
[![R-CMD-check](https://github.com/haliyan/fars.functions/workflows/R-CMD-check/badge.svg)](https://github.com/haliyan/fars.functions/actions)
<!-- badges: end -->

The goal of fars.functions is to read in, summarize, and map FARS (Fatality Analysis Reporting System) data on car accidents in the United States. This package was created as part of the Coursera course "Building R Packages" by JHU. 

### VERY IMPORTANT NOTES FOR PEER REVIEWERS

**PLEASE READ (2 NOTES BELOW)**

**NOTE 1**

**This package has a R-CMD-Check badge provided by GitHub Actions and not a Travis badge.**

Since the time of the creation of the course, Travis CI has changed significantly. It is no longer as straightforward as outlined in the course materials to use Travis for open-source projects. My understanding is that using their services for open-source projects such as this one requires signing up for a plan/trial (which in turn requires disclosing credit card information to Travis, which I wish to avoid if at all possible) and applying for open source credits. See Travis's statements on open source credits [here](https://docs.travis-ci.com/user/billing-faq/#what-if-i-am-building-open-source) and trial plan [here](https://docs.travis-ci.com/user/billing-overview/#free-trial-plan).

**Due to these issues and as per the suggestion of R, I have used GitHub Actions instead of Travis to check my package.**

You can verify that the log contains NO errors, warnings, or notes by going to the "Actions" tab in my GitHub repo and clicking on the TOP workflow run (it should have a blue check mark beside it). 

**I am asking that you please, please be considerate of this situation when grading my work!** 

**NOTE 2**

In order to pass the tests I set up in the tests sub directory, I had to make some changes to the original `fars_read` and `fars_summarize_years` functions. 

* Changing the `dplyr::tbl_df()` function (which has been deprecated) to the `tibble::as_tibble()` function in `fars.read`. 

* Adding the line `filename <- system.file("extdata", filename, package= "fars.functions"` to `fars.read`. 

* Changing `n()` to `dplyr::n()` in `fars_summarize_years`.

Omitting these changes causes problems with running examples and tests, building the vignette, creating the README file, and passing checks. **No other changes have been made to the original code.** 


## Installation

You can install the development version of `fars.functions` in R by entering `install_github("haliyan/fars.functions")` into the console. 

*Note: Using the* `install_github` *function requires that the* `devtools` *package be installed and loaded.* `devtools` *can be installed using* `install.packages("devtools")` *and loaded using* `library(devtools)`

Once the `fars.functions` package is installed, load it using `library(fars.functions)`. 

```{r,echo=FALSE}
library(fars.functions)
library(knitr)
```


**Examples**

The fars.functions package function `make_filename` creates a FARS file name of the correct format for a certain year supplied by the user. 
This format is `accident_X.csv.bz2`, where X is the user-supplied year:

```{r}
make_filename("2013")
```


The function `fars_read` reads in a FARS file of a certain file name and creates a tibble of the data. 
```{r}
fars_read("accident_2013.csv.bz2")
```

This function can also be used together with `make_filename` and the pipe operator `%>%` from the `magrittr` package, especially if the correct file name format is not known. 

```{r}
library(magrittr)
make_filename("2013") %>% 
  fars_read()
```


The `fars_read_years` takes a list or vector of years and produces a list of tibbles. Each tibble contains the months in one of the years in the supplied vector of years. The number of rows for each month represents the number of accidents that occurred in that month.

```{r}
fars_read_years(c("2013","2014"))
```


The `fars_summarize_years` function also takes a list or vector of years and produces a single tibble. The rows of the tibble represent the 12 months in a year. The columns of the tibble are the years that were supplied to the function. The values contained in the tibble are the numbers of accidents in a particular month of a particular year. 

```{r}
fars_summarize_years(c("2013","2014"))
```


The `fars_map_state` function creates a map of accidents in a given state in a given year. For example, the following code creates a map of the accidents in the state of Wisconsin in 2014:

```{r}
fars_map_state(55, 2014) 
```


*Note: for some state numbers, the* `fars_map_state` *function will not work, for various reasons. A quick summary table of these is included below. For more details plus a full table of all 56 state numbers and corresponding state names, please see the package vignette. You can instruct R to build the vignette when installing from GitHub by specifying* `build_vignettes=TRUE` *when installing with* `install_github`.

```{r, echo=FALSE}
data.frame(Numbers = c(2,3,7,14,15,43,52), 
           Reasons = c("Alaska; out of bounds", 
             "American Samoa; no data", 
             "Canal Zone; no data",
             "Guam; no data",
             "Hawaii; out of bounds", 
             "Puerto Rico; no data", 
             "Virgin Islands; no data")) %>% 
  kable()
```

**License**

GPL 
